\documentclass{article}
\usepackage{graphicx} % Required for inserting images
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{verbatim}
\usepackage{hyperref}
\usepackage{mathtools}
\usepackage{enumerate}

\usepackage{xy}
\xyoption{all}

\input{chillmacros}


\title{Applied Category Theory}
\author{Caleb Hill}
\date{Fall 2025}

\begin{document}
\maketitle


\section*{9/3/2025}

\subsection*{Why?}
Why should you care about studying the coming content and applying it to your field?

Physicists:
\begin{itemize}
    \item The particles in the standard model {\it are} irreducible representations. 
    So rep theory is crucial to you.
    \item Monoidal categories give a good framework for understanding QM.
\end{itemize}
Computer scientists:
\begin{itemize}
    \item It gives a framework for the Curry-Howard correspondence (proofs are programs).
\end{itemize}
Me:
\begin{itemize}
    \item Began as a study of ``analogies'' and turned into a study of nifty algebraic gadgets.
    \item It's written in a {\it wild} language, and learning languages is fun.
\end{itemize}


\subsection*{Plan}
The {\bf first goal} is to define monoidal categories with some context. 
The {\bf second goal} is to describe a ``skeletal'' category defined by diagrammatics.
To accomplish the first goal, we will study things including:
\begin{itemize}
    \item Algebraic objects (groups, vector spaces, ...) and maps between them 
    \item Subobjects, images, combining objects ($\times$, $\otimes$, $\oplus$, ...) 
    \item Categories ($Grp$, $Set$, $PoSet$, $\N$, $\Vecc$, ...)
\end{itemize}
I'd like to have as little fat on this as necessary.
That is, not get sidetracked studying, for instance, too much of the internal structure of these objects.
I want to give many examples and try to build intuition.
For the second goal we'll study things including:
\begin{itemize}
    \item Representations and maps ($T(gv) = gT(v)$)
    \item $\Rep(D_3)$ in detail
\end{itemize}
Up to this point I have a strong vision of where we're going. 
After this we can go where the interest steers us.

This plan is incomplete and non-exhaustive.




\subsection*{Groups}
This is the best onramp to categories I know of, so bear with me through some basics.
\begin{definition}
    A {\bf group} is a triple \[ (G,\mu,e)\] where $G$ is a set, 
    $\mu:G\times G\to G$ is a function, and $e\in G$, such that 
    \begin{equation*}\tag{Associativity}
        \forall a,b,c\in G, \quad \mu(\mu(a,b),c) = \mu(a,\mu(b,c))
    \end{equation*}
    \begin{equation*}\tag{Identity}
        \forall a\in G, \quad \mu(a,e) = \mu(e,a) = a
    \end{equation*}
    \begin{equation*}\tag{Inverse}
        \forall a\in G, \exists b\in G, \quad\text{such that}\quad \mu(a,b)=\mu(b,a)=e
    \end{equation*}
\end{definition}


\end{document}
